---
title: "R interface for GraphFrames"
output:
  github_document:
    fig_width: 9
    fig_height: 5
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(eval = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(fig.path = "tools/readme/", dev = "png")
```

[![Build Status](https://travis-ci.org/rstudio/graphframes.svg?branch=master)](https://travis-ci.org/rstudio/graphframes) [![Coverage status](https://codecov.io/gh/rstudio/graphframes/branch/master/graph/badge.svg)](https://codecov.io/github/rstudio/graphframes?branch=master) [![CRAN status](https://www.r-pkg.org/badges/version/graphframes)](https://cran.r-project.org/package=graphframes)

- Support for [GraphFrames](https://graphframes.github.io/) which aims to provide the functionality of [GraphX](http://spark.apache.org/graphx/).
- Perform graph algorithms like: [PageRank](https://graphframes.github.io/api/scala/index.html#org.graphframes.lib.PageRank), [ShortestPaths](https://graphframes.github.io/api/scala/index.html#org.graphframes.lib.ShortestPaths) and many [others](https://graphframes.github.io/api/scala/#package).
- Designed to work with [sparklyr](https://spark.rstudio.com) and the [sparklyr extensions](http://spark.rstudio.com/extensions.html).

## Installation

For those already using `sparklyr` simply run:

```{r eval=FALSE}
install.packages("graphframes")
# or, for the development version,
# devtools::install_github("rstudio/graphframes")
```

Otherwise, install first `sparklyr` from CRAN using:

```{r eval=FALSE}
install.packages("sparklyr")
```

The examples make use of the `highschool` dataset from the `ggplot` package.

## Getting Started

We will calculate [PageRank](https://en.wikipedia.org/wiki/PageRank) over the `highschool` dataset as follows:

```{r message=FALSE}
library(graphframes)
library(sparklyr)
library(dplyr)

# connect to spark using sparklyr
sc <- spark_connect(master = "local", version = "2.1.0")

# copy highschool dataset to spark
highschool_tbl <- copy_to(sc, ggraph::highschool, "highschool")

# create a table with unique vertices using dplyr
vertices_tbl <- sdf_bind_rows(
  highschool_tbl %>% distinct(from) %>% transmute(id = from),
  highschool_tbl %>% distinct(to) %>% transmute(id = to)
)

# create a table with <source, destination> edges
edges_tbl <- highschool_tbl %>% transmute(src = from, dst = to)

gf_graphframe(vertices_tbl, edges_tbl) %>%
  gf_pagerank(reset_prob = 0.15, max_iter = 10L, source_id = "1")
```

## Further Reading

Appart from calculating `PageRank` using `gf_pagerank`, the following functions are available: 

- gf_bfs: Breadth-first search (BFS).
- gf_connected_components: Connected components.
- gf_shortest_paths: Shortest paths algorithm.
- gf_scc: Strongly connected components.
- gf_triangle_count: Computes the number of triangles passing through each vertex and others.

For instance, one can calcualte the degrees of vertices using `gf_degrees` as follows:

```{r message=FALSE}
gf_graphframe(vertices_tbl, edges_tbl) %>% gf_degrees()
```

In order to visualize large `graphframe`s, one can use `sample_n` and then use `ggraph` with `igraph` to visualize the graph as follows:

```{r message=FALSE}
library(ggraph)
library(igraph)

graph <- highschool_tbl %>%
  sample_n(20) %>%
  collect() %>%
  graph_from_data_frame()

ggraph(graph, layout = 'kk') + 
    geom_edge_link(aes(colour = factor(year))) + 
    geom_node_point() + 
    ggtitle('An example')
```

Finally, we disconnect from Spark:

```{r}
spark_disconnect(sc)
```
